\normalfont{disCOVIDer19}

### 2020-09-17 Documentation/article_4.tex
PUB

In order to evaluate the effectiveness of the containment of the virus, we considered to use the Shannon Entropy:

@@

where, in our case, we considered $\textbf{P}(x_{i})$ as the number of currently infected people for a single cluster $i$ (the sum runs on the overall number of clusters). If for each cluster the ratio is the same, the containment of the virus between the clusters is failed: in this case a large value of entropy will be found; on the other hand if a strong dissimilarity between the clusters is present (e.g. the cases are located only in to a single one), the entropy will be low. The situation can be compared with the expansion of a gas: suppose that a toxic gas spreads in to building that has different rooms; the gas, due to the thermal noise, tends to expand uniformly into the whole building so that an homogeneous concentration will be obtained at least. This natural spread can be stopped with an external work, for instance by closing some doors or by building walls: in this case the spread will be stopped. It is worth nothing that this analogy, that motivated our use of the entropy index is valid as long as the virus is in its expansive phase (more precisely if the overall R(t) is $\geq $ 1): otherwise, differently with respect to the molecules of the gas that are constant over time, the active case reduces as the time increase, thus the analogy and so the index have no further meaning.

>>>>

In order to evaluate the effectiveness of the containment of the virus, we considered to use the Gibbs formulation of the entropy:

@@

where, in our case, we considered $\textbf{P}(x_{i})$ as the number of currently infected people for the single cluster $i$ (the sum runs on the overall number of clusters) and the the Boltzmann's constant was set to 1. Note that, in a different context, this expression is the Shannon Entropy where $\textbf{P}(x_{i})$ is the probability of the outcome $x_{i}$. If for each cluster the ratio is the same, the containment of the virus between the clusters is failed: in this case a large value of entropy will be found; on the other hand if a strong dissimilarity between the clusters is present (e.g. the cases are located only in to a single one), the entropy will be low. The situation can be compared with the expansion of a gas: suppose that a toxic gas spreads in to building that has different rooms; the gas, due to the thermal noise, tends to expand uniformly into the whole building so that an homogeneous concentration will be obtained at least. This behaviour is due to the fact that, since the atoms are indistinguishable, the configurations in which all the atoms are distributed isotropically in to the space are much more than the configurations in which the atoms are localized \cite{kittel1998thermal} .  This natural spread can be stopped with an external work, for instance by closing some doors or by building walls: in this case the spread will be stopped. It is worth nothing that this analogy, that motivated our use of the entropy index is valid as long as the virus is in its expansive phase (more precisely if the overall R(t) is $\geq $ 1): otherwise, differently with respect to the molecules of the gas that are constant over time, the active case reduces as the time increase, thus the analogy and so the index have no further meaning.
